### 2024-06-12 chap_atom_mol/chap_atom_mol.tex
Integration of comments in rebuttals

In a country like Belgium, reaching the goal of energy transition will not be a ``winner-takes-all'' situation but rather a combination of solutions \cite{Limpens2020,limpens2021generating}. However, this chapter focuses on this atom-molecules dilemma. Often compared, if not opposed, to local renewables like wind and solar \cite{suna2016nuclear,khatib2016economics}, this chapter rather assesses the integration of nuclear energy in the future versus the need to import renewable molecules from abroad, in a country where the local potential of \gls{VRES} is limited compared to its \gls{EUD}.

Section \ref{sec:atom_mol:results_deter} targets the impact of integrating \gls{SMR} from 2040 onward on the whole-energy system, in a deterministic way (\ie considering only nominal values of the parameters).

@@

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimise a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Section \ref{sec:cs:uncertainty}). Accounting for uncertainties presented in Section \ref{sec:meth:UQ}, Section \ref{sec:atom_mol:results_uq} will identify the key factors driving higher or lower imports of electrofuels as well as the installation of \gls{SMR}, by the end of the transition, \ie 2050.

Relying on nuclear energy for some is highly controversial. On top of purely techno-economic aspects, non-exhaustively mentioned beforehand, there are other ethical, societal or even political considerations to account for when addressing this question \cite{kempf2022}. In the same sense, on top of ethical or geopolitical aspects, one could question the availability of the imported electrofuels, assumed to be unlimited in this work (see Chapter \ref{chap:case_study}). In their work, \citet{lefebvre2022electrofuel} investigated this topic for the case of Belgium considering lower and upper bounds in terms of availability based on, on the one hand, the already signed agreements with the exporting countries and, on the other hand, their maximal technological potential of \gls{VRES}, respectively. Their lower bound resulted in a total availability of electrofuels that is one order of magnitude lower than the needs provided by the cost-optimum carbon-neutral Belgium in 2050. Like the rest of this thesis, the purpose here is only to expose the impact of integrating \gls{SMR} as well as the need of importing electrofuels in the Belgian energy landscape, on a strictly techno-economical point of view with a cost-based optimisation. This is why, for the sake of transparency, the model and the data are documented and openly available online \cite{readthedocs_pathway} and in the Appendices \ref{app:EnergyScope} and \ref{app:case_study}.

@@

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost. Where the cost of purchasing electrofuels is the top-1 parameter, \gls{SMR}-related parameters have a negligigble impact on this cost.}
\label{tab:UQ_short}
\centering
\begin{tabular}{l c c}
\toprule
\textbf{Parameter}  & \textbf{Ranking} & \textbf{Sobol' index} \\	
\midrule
\textbf{Purchase electrofuels} & 1 & 46.8\% \\
Industry EUD & 2 & 23.2\% \\
Interest rate & 3 & 12.0\% \\
Purchase fossil fuels  & 4 & 5.7\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{Potential capacity \gls{SMR}} & 11 & 0.9\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{CAPEX \gls{SMR}} & 33 & $<$0.1\% \\
\bottomrule

@@

As the driving factor for the annualisation and the salvage value of the assets, the interest rate has a 12\% Sobol' index. The model considers an overall interest rate for the entire system (\ie 1.5\% as a nominal value). In practice, the interest rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.

>>>>

In a country like Belgium, reaching the goal of energy transition will not be a ``winner-takes-all'' situation but rather a combination of solutions \cite{Limpens2020,limpens2021generating}. However, this chapter focuses on this atom-molecules dilemma. Often compared, if not opposed, to local renewables like wind and solar \cite{suna2016nuclear,khatib2016economics}, this chapter rather assesses the integration of nuclear energy in the future versus the need to import renewable molecules from abroad, in a country where the local potential of \gls{VRES} is limited compared to its \gls{EUD}. Relying on nuclear energy for some is highly controversial. On top of purely techno-economic aspects, non-exhaustively mentioned beforehand, there are other ethical, societal or even political considerations to account for when addressing this question \cite{kempf2022}. In the same sense, on top of ethical or geopolitical aspects, one could question the availability of the imported electrofuels, assumed to be unlimited in this work (see Chapter \ref{chap:case_study}). In their work, \citet{lefebvre2022electrofuel} investigated this topic for the case of Belgium considering lower and upper bounds in terms of availability based on, on the one hand, the already signed agreements with the exporting countries and, on the other hand, their maximal technological potential of \gls{VRES}, respectively. Their lower bound resulted in a total availability of electrofuels that is one order of magnitude lower than the needs provided by the cost-optimum carbon-neutral Belgium in 2050. Like the rest of this thesis, the purpose here is only to expose the impact of integrating \gls{SMR} as well as the need of importing electrofuels in the Belgian energy landscape, on a strictly techno-economical point of view with a cost-based optimisation. This is why, for the sake of transparency, the model and the data are documented and openly available online \cite{readthedocs_pathway} and in the Appendices \ref{app:EnergyScope} and \ref{app:case_study}.

Section \ref{sec:atom_mol:results_deter} targets the impact of integrating \gls{SMR} from 2040 onward on the whole-energy system, in a deterministic way (\ie considering only nominal values of the parameters). Accounting for uncertainties presented in Section \ref{sec:meth:UQ}, Section \ref{sec:atom_mol:results_uq} will identify the key factors driving higher or lower imports of electrofuels as well as the installation of \gls{SMR}, by the end of the transition, \ie 2050.

@@

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimise a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Section \ref{sec:cs:uncertainty}).

@@

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost. Where the cost of purchasing electrofuels is the top-1 parameter, \gls{SMR}-related parameters have a negligigble impact on this cost.}
\label{tab:UQ_short}
\centering
\begin{tabular}{l c c}
\toprule
\textbf{Parameter}  & \textbf{Ranking} & \textbf{Sobol' index} \\	
\midrule
\textbf{Purchase electrofuels} & 1 & 46.8\% \\
Industry EUD & 2 & 23.2\% \\
Discount rate & 3 & 12.0\% \\
Purchase fossil fuels  & 4 & 5.7\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{Potential capacity \gls{SMR}} & 11 & 0.9\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{CAPEX \gls{SMR}} & 33 & $<$0.1\% \\
\bottomrule

@@

As the driving factor for the annualisation and the salvage value of the assets, the discount rate has a 12\% Sobol' index. As detailed in Chapter \ref{chap:chap_methodo}, the model considers an overall discount rate for the entire system (\ie 1.5\% as a nominal value). In practice, the discount rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.

### 2024-06-12 chap_case_study/chap_case_study.tex
Integration of comments in rebuttals

This work considers nine groups of uncertain parameters: (i) the cost of purchasing imported energy carriers; (ii) the investment cost (\ie CAPEX) of some technologies, mostly related to the mobility sector and the integration of renewables; (iii) the maintenance cost (\ie OPEX) of every technology; (iv) the consumption of electric and fuel cells vehicles in the mobility sector; (v) the potential installed capacity of renewables; (vi) the hourly load factor of renewables accounting for variability of solar irradiance or wind speed; (vii) the availability of resources considered as limited (\ie biomass and electricity); (viii) the end-use-demands split per sector of activities (\ie households, services, passenger mobility and industry) and (ix) other parameters like the interest rate or the modal share change in different key sectors. A particular attention is to pay to the potential installation of \gls{SMR}, at the bottom of \Cref{tab:UC_full}. As detailed before, the commercial availability of such a technology is uncertain but would not be before 2040. Consequently, for \gls{SMR}, the parameter $f_{\mathrm{max,SMR}}$ influences the maximum capacity to install to translate somehow the readiness of this technology. Arbitrarily, we have then assumed the following probability of availability of such a technology: 10\% of chance to be installable from 2040, 20\% from 2045 and 40\% from 2050\footnote{In other words, if this parameter, ranging between 0 and 1, is (i) smaller than 0.6, there is no possibility to install \gls{SMR} during the transition; (ii) between 0.6 and 0.8, these 6~GW can be installed only in 2050; (iii) between 0.8 and 0.9, these can be installed from 2045 onward and; (iv) higher than 0.9, the prescribed maximum capacity can be installed from 2040 onward. }. Based on the local sensitivity analysis carried out by \citet{PATHS2050}, the current work also considers a [-40\%; +44\%] range on the CAPEX of SMR, on top of the uncertainty about its availability. Finally, the cost of purchasing renewable electrofuels presents a wide range, [-64.3\%; +179.8\%], like the other imported commodities.

@@

\multirow{6}{*}{\textbf{Miscellaneous}} &$i_{\mathrm{rate}}$  & Interest rate & I & -46.2\% & 46.2\% \\
& $\%_{\mathrm{pub,max}}$ & Max share of public transport & I & -10\% & 10\% \\
& $\Delta_{\mathrm{change,freight}}$ & Modal share change freight mobility & - & -30\% & 30\% \\
& $\Delta_{\mathrm{change,pass}}$ & Modal share change passenger mobility & - & -30\% & 30\% \\
& $\Delta_{\mathrm{change,LT\_heat}}$ & Modal share change LT-heat & - & -30\% & 30\% \\

@@

%\begin{table}[htbp!]
%\caption{Illustration of the uncertainty characterisation for different parameters for the year 2025. }
%\label{tab:UC_short}
%\begin{minipage}{\linewidth}
%\centering
%\resizebox{\textwidth}{!}{
%\begin{tabular}{l l l c c c}
%\toprule
%\multirow{2}{*}{\textbf{Category}} & \multirow{2}{*}{\textbf{Parameter}} & \multirow{2}{*}{\textbf{Meaning}} & \multirow{2}{*}{\textbf{Type}\footnote{\label{foot:type_uncert_range}Per \citet{Moret2017PhDThesis}, \og I: investment-type, II: operation-type (constant uncertainty over time), III: operation-type (uncertainty increasing over time)\fg. }}  & \multicolumn{2}{c}{\textbf{Relative variation\footnote{\label{foot:nom_val_uncert}The nominal values for each of the parameters is 0, meaning no variation compared to the nominal values of the impacted parameter in the model. }}}\\
%    & & & &	 min 	&	 max \\ 	
%\midrule		
%\multirow{2}{*}{\textbf{Cost of purchasing}} & $c_{\mathrm{op,fossil}}$ & Purchase fossil fuels & II & -64.3\% & 179.8\% \\
%& $c_{\mathrm{op,electrofuels}}$ & Purchase electrofuels & II & -64.3\% & 179.8\% \\
%\midrule
%\multirow{5}{*}{\textbf{Investment cost}} &$c_{\mathrm{inv,car}}$ & CAPEX car  & I & -21.6\% & 25.0\% \\
%& $c_{\mathrm{inv,e\_prop}}$ & CAPEX electric motor & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,fc\_prop}}$ & CAPEX fuel cell engine & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,PV}}$ & CAPEX PV & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,nuclear\_SMR}}$ & CAPEX \gls{SMR}\footnote{\label{foot:range_SMR}This range has been inferred from the local sensitivity analysis performed by \citet{PATHS2050}.}& I & -40.0\% & 44.0\% \\
%\midrule
%\multirow{1}{*}{\textbf{Consumption}} &$\eta_{\mathrm{e\_prop}}$ & Consumption electric vehicles & I & -28.7\% & 28.7\% \\
%\midrule
%\multirow{2}{*}{\textbf{Potential installed capacity}} &$f_{\mathrm{max,PV}}$ & Max capacity PV & I & -24.1\% & 24.1\% \\
%& $f_{\mathrm{max,windon}}$ & Max capacity onshore wind & I & -24.1\% & 24.1\% \\
%\midrule
%\multirow{2}{*}{\textbf{Hourly load factor}} & $c_{\mathrm{p,t,PV}}$ & Hourly load factor PV & II & -22.1\% & 22.1\% \\
%& $c_{\mathrm{p,t,winds}}$ & Hourly load factor wind turbines & II & -22.1\% & 22.1\% \\
%\midrule
%\multirow{2}{*}{\textbf{Resource availability}} & $avail_{\mathrm{elec}}$ & Available electricity import & I & -32.1\% & 32.1\% \\
%& $avail_{\mathrm{biomass}}$ & Available local biomass & I & -32.1\% & 32.1\% \\
%\midrule
%
%\multirow{2}{*}{\textbf{End-use demand}} & $pass\_EUD$ & Passenger mobility EUD & III & -7.5\% & 7.5\% \\
%& $industry\_EUD$ & Industry EUD & III & -20.5\% & 16.0\% \\
%\midrule
%
%\multirow{4}{*}{\textbf{Miscellaneous}} &$i_{\mathrm{rate}}$  & Interest rate & I & -46.2\% & 46.2\% \\
%& $\Delta_{\mathrm{change,freight}}$ & Modal share change freight mobility & - & -30\% & 30\% \\
%& $\Delta_{\mathrm{change,pass}}$ & Modal share change passenger mobility & - & -30\% & 30\% \\
%& $f_{\mathrm{max,SMR}}$ & Potential capacity \gls{SMR} & - & 0 & 1 \\
%
%\bottomrule							
%
%\end{tabular}}
%\end{minipage}
%\end{table}

>>>>

This work considers nine groups of uncertain parameters: (i) the cost of purchasing imported energy carriers; (ii) the investment cost (\ie CAPEX) of some technologies, mostly related to the mobility sector and the integration of renewables; (iii) the maintenance cost (\ie OPEX) of every technology; (iv) the consumption of electric and fuel cells vehicles in the mobility sector; (v) the potential installed capacity of renewables; (vi) the hourly load factor of renewables accounting for variability of solar irradiance or wind speed; (vii) the availability of resources considered as limited (\ie biomass and electricity); (viii) the end-use-demands split per sector of activities (\ie households, services, passenger mobility and industry) and (ix) other parameters like the discount rate or the modal share change in different key sectors. A particular attention is to pay to the potential installation of \gls{SMR}, at the bottom of \Cref{tab:UC_full}. As detailed before, the commercial availability of such a technology is uncertain but would not be before 2040. Consequently, for \gls{SMR}, the parameter $f_{\mathrm{max,SMR}}$ influences the maximum capacity to install to translate somehow the readiness of this technology. Arbitrarily, we have then assumed the following probability of availability of such a technology: 10\% of chance to be installable from 2040, 20\% from 2045 and 40\% from 2050\footnote{In other words, if this parameter, ranging between 0 and 1, is (i) smaller than 0.6, there is no possibility to install \gls{SMR} during the transition; (ii) between 0.6 and 0.8, these 6~GW can be installed only in 2050; (iii) between 0.8 and 0.9, these can be installed from 2045 onward and; (iv) higher than 0.9, the prescribed maximum capacity can be installed from 2040 onward. }. Based on the local sensitivity analysis carried out by \citet{PATHS2050}, the current work also considers a [-40\%; +44\%] range on the CAPEX of SMR, on top of the uncertainty about its availability. Finally, the cost of purchasing renewable electrofuels presents a wide range, [-64.3\%; +179.8\%], like the other imported commodities.

@@

\multirow{6}{*}{\textbf{Miscellaneous}} &$i_{\mathrm{rate}}$  & Discount rate & I & -46.2\% & 46.2\% \\
& $\%_{\mathrm{pub,max}}$ & Max share of public transport & I & -10\% & 10\% \\
& $\Delta_{\mathrm{change,freight}}$ & Modal share change freight mobility & - & -30\% & 30\% \\
& $\Delta_{\mathrm{change,pass}}$ & Modal share change passenger mobility & - & -30\% & 30\% \\
& $\Delta_{\mathrm{change,LT\_heat}}$ & Modal share change LT-heat & - & -30\% & 30\% \\

@@

%\begin{table}[htbp!]
%\caption{Illustration of the uncertainty characterisation for different parameters for the year 2025. }
%\label{tab:UC_short}
%\begin{minipage}{\linewidth}
%\centering
%\resizebox{\textwidth}{!}{
%\begin{tabular}{l l l c c c}
%\toprule
%\multirow{2}{*}{\textbf{Category}} & \multirow{2}{*}{\textbf{Parameter}} & \multirow{2}{*}{\textbf{Meaning}} & \multirow{2}{*}{\textbf{Type}\footnote{\label{foot:type_uncert_range}Per \citet{Moret2017PhDThesis}, \og I: investment-type, II: operation-type (constant uncertainty over time), III: operation-type (uncertainty increasing over time)\fg. }}  & \multicolumn{2}{c}{\textbf{Relative variation\footnote{\label{foot:nom_val_uncert}The nominal values for each of the parameters is 0, meaning no variation compared to the nominal values of the impacted parameter in the model. }}}\\
%    & & & &	 min 	&	 max \\ 	
%\midrule		
%\multirow{2}{*}{\textbf{Cost of purchasing}} & $c_{\mathrm{op,fossil}}$ & Purchase fossil fuels & II & -64.3\% & 179.8\% \\
%& $c_{\mathrm{op,electrofuels}}$ & Purchase electrofuels & II & -64.3\% & 179.8\% \\
%\midrule
%\multirow{5}{*}{\textbf{Investment cost}} &$c_{\mathrm{inv,car}}$ & CAPEX car  & I & -21.6\% & 25.0\% \\
%& $c_{\mathrm{inv,e\_prop}}$ & CAPEX electric motor & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,fc\_prop}}$ & CAPEX fuel cell engine & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,PV}}$ & CAPEX PV & I & -39.6\% & 39.6\% \\
%& $c_{\mathrm{inv,nuclear\_SMR}}$ & CAPEX \gls{SMR}\footnote{\label{foot:range_SMR}This range has been inferred from the local sensitivity analysis performed by \citet{PATHS2050}.}& I & -40.0\% & 44.0\% \\
%\midrule
%\multirow{1}{*}{\textbf{Consumption}} &$\eta_{\mathrm{e\_prop}}$ & Consumption electric vehicles & I & -28.7\% & 28.7\% \\
%\midrule
%\multirow{2}{*}{\textbf{Potential installed capacity}} &$f_{\mathrm{max,PV}}$ & Max capacity PV & I & -24.1\% & 24.1\% \\
%& $f_{\mathrm{max,windon}}$ & Max capacity onshore wind & I & -24.1\% & 24.1\% \\
%\midrule
%\multirow{2}{*}{\textbf{Hourly load factor}} & $c_{\mathrm{p,t,PV}}$ & Hourly load factor PV & II & -22.1\% & 22.1\% \\
%& $c_{\mathrm{p,t,winds}}$ & Hourly load factor wind turbines & II & -22.1\% & 22.1\% \\
%\midrule
%\multirow{2}{*}{\textbf{Resource availability}} & $avail_{\mathrm{elec}}$ & Available electricity import & I & -32.1\% & 32.1\% \\
%& $avail_{\mathrm{biomass}}$ & Available local biomass & I & -32.1\% & 32.1\% \\
%\midrule
%
%\multirow{2}{*}{\textbf{End-use demand}} & $pass\_EUD$ & Passenger mobility EUD & III & -7.5\% & 7.5\% \\
%& $industry\_EUD$ & Industry EUD & III & -20.5\% & 16.0\% \\
%\midrule
%
%\multirow{4}{*}{\textbf{Miscellaneous}} &$i_{\mathrm{rate}}$  & Discount rate & I & -46.2\% & 46.2\% \\
%& $\Delta_{\mathrm{change,freight}}$ & Modal share change freight mobility & - & -30\% & 30\% \\
%& $\Delta_{\mathrm{change,pass}}$ & Modal share change passenger mobility & - & -30\% & 30\% \\
%& $f_{\mathrm{max,SMR}}$ & Potential capacity \gls{SMR} & - & 0 & 1 \\
%
%\bottomrule							
%
%\end{tabular}}
%\end{minipage}
%\end{table}

### 2024-06-12 chap_intro_ccl/conclusion.tex
Integration of comments in rebuttals

% THE RESULTS WE GOT (WHAT?)
Given the \ce{CO2}-budget to stick to a +1.5$^{\circ}$C increase of global temperature and the presumably high levels of demands,  aiming at a cost-optimum transition, Belgium should first opt for more efficiency to reduce by one third the primary energy consumption. Then comes the full deployment of local solar and wind energies that would represent about 30\% of the mix by 2050. Due to the limited potential of local renewables, electrofuels are found to be the third pillar of the Belgian transition. The uncertainty on the total cost of the transition is mainly due to the uncertainty on the cost of purchasing electrofuels (47\%), on the industrial end-use demand (23\%), on the interest rate (12\%) and on the price of fossil fuels (6\%).

>>>>

% THE RESULTS WE GOT (WHAT?)
Given the \ce{CO2}-budget to stick to a +1.5$^{\circ}$C increase of global temperature and the presumably high levels of demands,  aiming at a cost-optimum transition, Belgium should first opt for more efficiency to reduce by one third the primary energy consumption. Then comes the full deployment of local solar and wind energies that would represent about 30\% of the mix by 2050. Due to the limited potential of local renewables, electrofuels are found to be the third pillar of the Belgian transition. The uncertainty on the total cost of the transition is mainly due to the uncertainty on the cost of purchasing electrofuels (47\%), on the industrial end-use demand (23\%), on the discount rate (12\%) and on the price of fossil fuels (6\%).

### 2024-06-12 chap_intro_ccl/intro.tex
Integration of comments in rebuttals

Despite the growing echo it finds in the scientific community \cite{o2018good}, sufficiency is lacking in the RED III. Explicitly mentioned by the IPCC for the first time in 2022 \cite{IPCC2022}, it is defined as ``\emph{a strategy for reducing, in absolute terms, the consumption and production of end-use products and services through changes in social practices in order to comply with environmental sustainability while ensuring an adequate social foundation for all people}'' \cite{lage2023citizens}. Due to its intrinsically political nature, studying sufficiency requires an interdisciplinary approach involving economy, sociology, psychology and political science, and is outside the scope and knowledge of engineering research \cite{schmidt2015interdisciplinary}. One of the objectives of this work is to support interdisciplinary projects in the assessment of sufficiency policy. \\

@@

Chapter \ref{chap:chap_RobPol} assesses the robustness of different roadmaps resulting from deterministic perfect foresight optimisation under certain conditions: REF, SMR and ROB. The first one, the reference case, considers nominal values for all the uncertain parameters. The SMR case is the one introduced in Chapter \ref{chap:atom_mol} where we allow the model to install \gls{SMR} from 2040 onward. Eventually, the ROB case accounts for the highest values of parameters having the biggest impact on the total cost of transition (\ie cost of purchasing energy carriers, industrial \gls{EUD}, interest rate and variable \gls{OPEX} of technologies).

>>>>

Despite the growing echo it finds in the scientific community \cite{o2018good}, sufficiency is lacking in the RED III. Explicitly mentioned by the IPCC for the first time in 2022 \cite{IPCC2022}, it is defined as ``\emph{a strategy for reducing, in absolute terms, the consumption and production of end-use products and services through changes in social practices in order to comply with environmental sustainability while ensuring an adequate social foundation for all people}'' \cite{lage2023citizens}. Due to its intrinsically political nature, studying sufficiency requires an interdisciplinary approach involving economy, sociology, psychology and political science, and is beyond the scope and knowledge of engineering research \cite{schmidt2015interdisciplinary}. Among all the lenses through which it is necessary to assess sufficiency policies, one of the objectives of this work is to support these interdisciplinary projects by providing informed techno-economic guidelines.\\

@@

Chapter \ref{chap:chap_RobPol} assesses the robustness of different roadmaps resulting from deterministic perfect foresight optimisation under certain conditions: REF, SMR and ROB. The first one, the reference case, considers nominal values for all the uncertain parameters. The SMR case is the one introduced in Chapter \ref{chap:atom_mol} where we allow the model to install \gls{SMR} from 2040 onward. Eventually, the ROB case accounts for the highest values of parameters having the biggest impact on the total cost of transition (\ie cost of purchasing energy carriers, industrial \gls{EUD}, discount rate and variable \gls{OPEX} of technologies).
